Report on Late Winter2013 Production: Image Differencing

### 2013-03-25 W13report.tex
Wordsmithing

We assume that science image $S(x,y)$ can be modeled as a convolution
of the template image $T(x,y)$ by a PSF--matching kernel $K(u,v,x,y)$.
We further assume that the kernel may be decomposed using a set of
basis functions $K(u,v) = \sum_i a_i K_i(u,v)$, where the coefficients
in front of each basis are determined through:
%
\begin{eqnarray}
C_i & \equiv & (K_i \otimes T) \\ \nonumber
b_{i}  & = & \sum_{x,y} {{C_i(x,y) S(x,y)}\over{\sigma^2(x,y)}}   \nonumber \\ 
M_{ij} & = & \sum_{x,y} {{C_i(x,y) C_j(x,y)}\over{\sigma^2(x,y)}}  \nonumber \\ 
a_{i}  & = & M^{-1}_{ij} b_{j} \nonumber 
\label{eq-soln}
\end{eqnarray}
where $\sigma^2(x,y)$ is the per--pixel variance stored in the {\tt
  variance} plane of each LSST {\tt exposure}.  To generate a
spatially varying model for the kernel, we assume that the relative
weights of the basis functions $a_i$ themselves vary spatially,
i.e. $K(u,v,x,y) = \sum_i a_i(x,y) K_i(u,v)$.  We also assume that
there is a differential background between the two images $b(x,y)$
that may be fit for using a low--order Chebyshev.  The image
difference is then calculated through $D(x,y) = S(x,y) - T(x,y)
\otimes K(u,v,x,y) - b(x,y)$.

The basis functions themselves $K_i(u,v)$ are a degree of freedom in
this problem.  The original implementations \citep{Alard98,Alard00}
used a set of 3 Gaussians (represented by our {\tt Config} variable
nGauss), each with a different sigma (sigGauss), and each modified by
a Laguerre polynomial to a given order (degGauss).  Subsequent studies
\citep[e.g.][]{2007AN....328...16I} have suggested that a constant
ratio be maintained between the different Gaussian widths, such that
$\sigma_{i+1} = \beta \times \sigma_{i}$.  We use the value $\beta =
2.0$ for this production.  We set the scale for sigma by noting that,
under the assumption that the Psfs of the images are Gaussian
($\sigma_S$ for the science image and $\sigma_T$ for the template
image), the sigma of the matching kernel should be simply $\sigma_K^2
= \sigma_S^2 - \sigma_T^2$.  We use this width for the central
Gaussian in our sequence, when we use more than one Gaussian, or the
sole Gaussian in the sequence when we use only one.

@@

In this production, we investigate two orders of operations for the
optimal filtering by the Psf.  In the first, we Psf--match the
template image to a pre--filtered science image:
\begin{eqnarray}
D_{Pre}(x,y) & = & S(x,y) \circ Psf_S(x,y,u,v) - T(x,y) \otimes K(u,v,x,y) \nonumber \\ 
D'_{Pre}(x,y) & = & D_{Pre}(x,y). \nonumber 
\end{eqnarray}
By pre--smoothing the image that we are matching the template to --
$S(x,y) \circ Psf_S(x,y,u,v)$ -- we more frequently avoid the case
of psf--matching {\it deconvolution}, where the FWHM of $T$ is
narrower than $S$.  This pre--filtering increases the FWHM of the
image by $\sqrt{2}$.  The Psf matching kernels will need to be
correspondingly larger to account for the larger Psfs.  We are able to
run detection directly on this image.

In the second, which is the classical implementation of the problem,
we create a difference image that is then post--filtered with the Psf
for detection:
\begin{eqnarray}
D_{Post}(x,y) & = & S(x,y) - T(x,y) \otimes K(u,v,x,y) \nonumber \\ 
D'_{Post}(x,y) & = & D_{Post}(x,y) \circ Psf_S(x,y,u,v).  \nonumber 
\end{eqnarray}

@@

Images $S(x,y)$ (optionally pre-filtered with $Psf_S(x,y,u,v)$) and
$T(x,y)$ are sent to ImagePsfMatchTask, whose purpose is to fit for
$K(u,v,x,y)$ and to produce $D(x,y)$.  The first step is to use the
Psf Gaussian sigmas to define the {\it sizes} of the Gaussians shapes
in $K_i(u,v)$.  As mentioned above, the main width is chosen to be
$\sqrt{\sigma_S^2 - \sigma_T^2}$.  We use variations on the
configuration where we use one Gaussian, with the above width, and
three Gaussians, the central of which has this width, and the others a
factor of $\beta = 2.0$ smaller and larger.  We modify each of the
Gaussians by a set of Laguerre polynomials of a given order.  The
smallest Gaussian is modified to the highest order, with the others
modified by order//2.  Figure~\ref{basis} provides an example basis
set with this configuration.  The total number of shapes in the basis
are: $\sum_i^{\rm nGauss} ({\rm degGauss}_i+1)\times({\rm
  degGauss}_i+2)/2$ \footnote{ A notable exception to this algorithm
  is found in the case where $\sigma_T > \sigma_S$, i.e. a sharpening
  must be done to the template to match it to the science image.  This
  is the case of deconvolution.  A theoretical prescription for basis
  design in this case is given in \cite{0266-5611-26-8-085002}.  In
  this algorithm, nGauss is {\it always} fixed to value 3, and
  degGauss is always fixed to 3.  The widths of the Gaussians are
  determined through the algorithm specified in
  \cite{0266-5611-26-8-085002}, using as inputs the sequence of
  Gaussians that we would have used to match a Gaussian of width
  $\sigma_S$ to $\sigma_T$ (i.e. as if we would have convolved the
  science image and not the template image).  There are a variety of
  reasons that actually operating on the science image is undesirable,
  including: the science image is lower S/N than the template; the
  science image is likely to have masked defects, including cosmic
  rays and bad pixels, whose reach would be expanded due to the
  convolution; unmasked features such as faint cosmic rays are more
  likely to look Psf--like after convolution (cosmic rays will likely
  have been sigma clipped out of the coadd).  However, it is not clear
  that deconvolving the template image is in principle any better.  A
  far better option, which we explore here, is to pre--filter the
  science image, which will in the vast majority cases provide a
  larger Psf (larger by $\sqrt{2}$) to match the template to.  This is
  the case for v6866601 in this study, where the postfiltered analysis
  yields deconvolution of the template, while the prefiltered analysis
  does not.
}.
%
The dimensions of the Psf matching kernel are
chosen to be 6 times the largest Gaussian width, with a minimum
kernelSize of 21x21 pixels.  For kernels that are significantly
smaller than this, the Gaussians have significant (non--zero) power at
the kernel boundaries, leading to square systematic artifacts at the
scale of the kernel in the difference images.

@@

For each source $j$ returned by DiaCatalogSourceSelector, we create a
{\tt KernelCandidate} instance, which holds substamps of $T_j(x,y)$
and $S_j(x,y)$ centered on the source.  These will be used to fit for
a local solution $K(u,v)$; the ensemble of local solutions at each
source's $x,y$ will be used to fit for the full spatial solution
$K(u,v,x,y)$.  We emphasize that the dimensions of the substamps is
important.  Note from equation~\ref{eq-soln} that the coefficients
$a_i$ are derived from a convolution of the template substamp with the
kernel basis functions $K_i$.  This convolution means that
kernelSize//2 pixels along each substamp edge are rendered unusable.
To maintain a significant number of pixels in $C_i$, we set the
KernelCandidate dimensions to be twice that of the kernel (i.e. the
number of pixels remaining in $C_i$ is equal to the number of pixels
in the kernel, which is fewer than the number of bases $i$).  For each
KernelCandidate, we solve for $K(u,v)$ and create a local difference
image $D_j(x,y)$.  We evaluate several statistics on this local
difference image, normalized by the square root of its variance, which
puts the pixels in units of sigma.  We measure the mean sigma, the RMS
of this distribution, the $\chi^2$ of these pixels normalized by the
degrees of freedom (number of pixels in $D(x,y)_j$ minus the number of
kernel bases), and the mean squared error.  These are referred to as
the LOCAL kernel metrics, and reflect how appropriate the chosen
kernel basis is.

@@

Finally, we run DipoleMeasurementTask on the detections to create {\tt
  DiaSources}.  DipoleMeasurementTask is an extension of
SourceMeasurementTask that includes specific dipole measurement
algorithms.  Measurement is necessarily different in the pre--filtered
and post--filtered data.  In particular, in post--filtering, detection
happens on the filtered maximum likelihood image and source
measurement on the associated pixels in the unfiltered difference
image.  In pre--filtering, the latter does not exist, and thus
detection and measurement happen on the same image.  This restricts
the types of measurements that may be made on the sources.  In
particular, the Psf flux is merely the fitted amplitude of the peak in
the image, and the Psf centroid its center.  While these measurements
are trivial to implement, additional work needs to be done on how to
implement the remainder of the measurement suite in a manner comparable
to the measurements on post--filtered data.

All {\tt DiaSources} are associated with the {\tt calexp's} {\tt
  sources} and with the reference catalog, with a permissive matching
radius of 3''.  {\tt DiaSources} that match with {\tt sources} may be
true residuals around stars {\it or} true false detections that are
also found in {\tt calexp} SourceDetectionTask.  For this reason, we
use the associations with the reference catalog objects as the
assessment if a given {\tt DiaSource} matches with a {\tt source} or
whether it is an orphan (e.g. a statistical fluctuation in the
background).  In addition, all {\tt KernelCandidates} are persisted
with the above {\tt LOCAL} and {\tt SPATIAL} metrics.  Repositories
are accessed with a specially written RepositoryIterator that
aggregates {\tt KernelCandidates} from a given {\tt source} across
multiple output repositories.

@@

We ran several pre--productions to tune: the dimensions of the kernels
and KernelCandidates; the widths of the Gaussians (sigGauss) based on
the FWHM of the input images; and the scaling of the Gaussians
($\beta$) in the basis sequence.  The main production runs included:
an outer loop over the use of pre--filtering or post--filtering for
detection; variations of the complexity of the Kernel basis, including
the number of Gaussians (nGauss = 1,3) and the degree of the modifying
Laguerre polynomials (degGauss = 1..6); and the spatial order of the
Chebyshev polynomials on $a_i(x,y)$ (spatialKernelOrder = 1..6).
Finally, we ran several post--productions focused on perturbing the
solutions that yielded the fewest numbers of DiaSources along specific
dimensions.  This included: the RMS of the WCS solution used by
RegisterTask; including a DC offset in the source positions prior to
RegisterTask, to determine the ability of the kernel basis to
accommodate astrometric misregistration; the size of the SpatialCells
used in the spatial interpolation, which effectively regulates the
numbers of constraints that go into the spatial fit; and modifying the
detection thresholds to determine how the number of false detections
scales with threshold.

@@

In order to investigate the bias--variance trade--off in the overall
fit, we calculate the MSE of each KernelCandidate's difference image
derived from the evaluation of the spatial model at that location.  We
define the bias as $\left| data - model \right|$, the variance as
$\left| (data - model)^2 \right|$, and the MSE as {\tt bias$^2$ +
  variance}.  In this context, the bias is the mean of the difference
image, and the variance is the mean square of the difference image.
In both cases, we normalize by the square root of the variance so that
pixels are in units of sigma.

@@

After running an ensemble of 144 pipeline configurations, we
aggregated the numbers of false detections.  We report the results for
the each of the 3 visits, and for pre-- and postfiltering for
detection.  We cut on {\tt flags.pixel.edge} to remove spurious
detections on the immediate border of the image; other than that we do
no other cuts other than requiring that the {\tt centroid.sdss} is
finite.  We count the DiaSources that match with the reference catalog
(3'' matching radius) via the {\tt refMatchId} value.  We find no
significant correlation of false detections with sensor within the
raft; therefore we report the aggregate results for all 9 sensors in
the analysis that follows.

@@

The numbers of false detections in each best case are given in
Table~\ref{tab-bestfp10}.  We undertook a manual inspection of all
difference images to associate the DiaSources with Sources in the
images, and also a 3'' association with the input source catalog.
These associations are listed in the N$_{Match}$ column of
Table~\ref{tab-bestfp10}, with the unmatched false positives in the
N$_{Orphan}$ column.  The former are likely to come from systematics
in the subtractions of stars, while the latter would be due to
fluctuations in the background.  We note 2 effects that will impact
N$_{Match}$.  The first is that, with a 3'' matching radius, 4\% of
random fluctuations will by--chance associate with a Source.  Assuming
that N$_{Orphan}$ represents 96\% of the background fluctuation
population, we estimate the number of fluctuations from that same
population that would have associated with Sources as N$_{Random}$.
Finally, we count all matches that occur within 100 pixels of the
image boundary, where we might expect the spatial extrapolation to
yield systematic residuals, as N$_{D<100}$.  To the extent that
(N$_{Random}$ + N$_{D<100}$) < N$_{Match}$, we have a small residual
set of systematic false detections in our imaging ensemble.  Aside
from the deconvolution run, this is at the level of 10 total false
detections over 9 sensors in 5 production runs, or 1 every 4--5
images.

@@

After finding the configurations that yielded the minimal numbers of
false detections, we then perturbed the solutions to examine how the
numbers of false detections scaled with different effects.  This
included, in order of importance for this production: the difference
image detection threshold; the number of KernelCandidates going into
the spatial model; and astrometric registration errors.  We also
undertake a theoretical analysis of the number of false detections
expected in a 4k x 4k image as a function of detection threshold, and
examine the requirements this sets on our ability to model the
variance and backgrounds in the images.  We compare the empirical
variation of false detections with detection threshold, and show that
our data are consistent with this.

@@

Because these images will have gone through multiple convolutions
before reaching the detection stage, it is expected that the
propagated per-pixel variance will not provide an exact representation
of the empirical variance in the image planes.  Because the
astrometric resampling uses a warping kernel that is designed to
preserve the noise properties of the images, we expect that the kernel
convolution and Psf--filtering will be the largest sources of error.

@@

We further investigate the variance properties of the difference
images immediately preceding detection, i.e. the difference image
itself in the case of prefiltering, and the difference image
correlated with its Psf in postfiltering.  These results are
summarized in Table~\ref{tab-variance2}.  We find that, empirically,
the images have very similar noise properties at the detection stage,
with the deconvolution visit having larger variance by $1\%$ but the
other visits having essentially equal noise properties, with a bias
towards the postfiltered data having slightly higher variance (while
the RMS is small, all values are greater than 1.0).  The medians of
the variance planes tell a different story though.  For the
deconvolution visit, the median per--pixel variance is $70\%$ higher.
While this will suppress the detections of false positives, it will
also significantly suppress the detection efficiency for true
variability.  For the other visits, the prefiltered per--pixel
variance is relatively {\it higher} by $3-4\%$.  This will decrease
the sensitivity of the prefiltered data to false positives; as shown
above, the rate is indeed smaller using the prefiltering pipeline.
This analysis demonstrates that this is primarily due to an {\it
  underestimate} of the per--pixel variance in the postfilter
processing.  As outlined by \cite{Price-Stacking}, this may be
accounted for by differences in pixel covariance when undergoing
multiple convolutions, which are not currently tracked.  We have not
yet definitively demonstrated that this is the case here.

@@

\item When deconvolving, the postfiltered variance is 70\% {\it
  higher} at the time of detection compared to the prefiltered data
  (Table~\ref{tab-variance2}).  This is consistent with the
  understanding that deconvolution increases high frequency noise in
  the images, and provides a {\it lower} detection efficiency in
  deconvolved data.  Prefiltering is clearly preferred in the case of
  deconvolution.

@@

\begin{figure}
\includegraphics[width=0.8\textwidth]{figures/shift.eps} \\
\caption{Figure illustrating the ability of the kernel basis to
  compensate, locally, for the effects of misregistration.  Each
  element of the grid shows a single KernelCandidate local difference
  image (visit=6866603, doPreFilter=True).  Columns indicate the
  Gaussian sigma of the basis, which is designed to have one Gaussian
  with the prescribed width, modified to fourth order.  Rows indicate
  the amplitude of the astrometric offset that was introduced into
  the Source positions before image-to-image registration.  
  In the {\it top} row, we see that even with a small astrometric
  offset, the sigma=2 basis is unable to produce a quality
  subtraction, because the width is inappropriate for matching the two
  input Psfs (the theoretical Gaussian sigma here is 3.4 pixels).  At
  sigma=4 the basis is able to produce a quality subtraction, and by
  sigma=5 the basis is too large to match the Psfs.  In {\it row 4},
  we show that even for an astrometric offset of 4 pixels (0.8''), the
  sigma=4 basis can produce a reasonable difference image without
  structured residuals.  However, by {\it row 5}, with an offset of 6
  pixels, none of the Gaussians provide a successful local difference
  image. 
  The ability of the basis to compensate for bulk astrometric offsets
  is a function of the kernel width, and the kernel width is itself a
  function of the two input Psf FWHMs.  Thus there is a complicated
  dependence of the ability of the kernel to shift the centroids of
  stars.  Roughly, the basis may compensate for local astrometric residuals
  at a scale up to $\sqrt{\sigma_S^2 - \sigma_T^2}$.
}
\label{kernel_offsets}
\end{figure}

>>>>

We assume that science image $S(x,y)$ can be modeled as a convolution
of the template image $T(x,y)$ by a PSF--matching kernel $K(u,v,x,y)$.
We further assume that the kernel may be decomposed using a set of
basis functions $K(u,v) = \sum_i a_i K_i(u,v)$, where the coefficients
in front of each basis are determined through:
%
\begin{eqnarray}
C_i & \equiv & (K_i \otimes T) \\ \nonumber
b_{i}  & = & \sum_{x,y} {{C_i(x,y) S(x,y)}\over{\sigma^2(x,y)}}   \nonumber \\ 
M_{ij} & = & \sum_{x,y} {{C_i(x,y) C_j(x,y)}\over{\sigma^2(x,y)}}  \nonumber \\ 
a_{i}  & = & M^{-1}_{ij} b_{j} \nonumber 
\label{eq-soln}
\end{eqnarray}
where $\sigma^2(x,y)$ is the per--pixel variance stored in the {\tt
  variance} plane of each LSST {\tt exposure}.  To generate a
spatially varying model for the kernel, we assume that the relative
weights of the basis functions $a_i$ themselves vary spatially,
i.e. $K(u,v,x,y) = \sum_i a_i(x,y) K_i(u,v)$.  We also assume that
there is a differential background between the two images $b(x,y)$
that may be fit for using a low--order Chebyshev polynomial.  The
image difference is then calculated through $D(x,y) = S(x,y) - T(x,y)
\otimes K(u,v,x,y) - b(x,y)$.

The basis functions themselves $K_i(u,v)$ are a degree of freedom in
this problem.  The original implementations \citep{Alard98,Alard00}
used a set of 3 Gaussians (represented by our {\tt Config} variable
nGauss), each with a different sigma (sigGauss), and each modified by
a Laguerre polynomial to a given order (degGauss).  Subsequent studies
\citep[e.g.][]{2007AN....328...16I} have suggested that a constant
ratio be maintained between the different Gaussian widths, such that
$\sigma_{i+1} = \beta \times \sigma_{i}$.  We use the value $\beta =
2.0$ for this production.  We set the overall scale for sigma by
noting that, under the assumption that the Psfs of the images are
Gaussian ($\sigma_S$ for the science image and $\sigma_T$ for the
template image), the sigma of the matching kernel should be simply
$\sigma_K^2 = \sigma_S^2 - \sigma_T^2$.  We use this width for the
central Gaussian in the basis sequences where we use more than one
Gaussian, or the sole Gaussian in the sequences when we use only one.

@@

In this production, we investigated two orders of operations for the
optimal filtering by the Psf.  In the first method, called
prefiltering, we Psf--match the template image to a pre--filtered
science image:
\begin{eqnarray}
D_{Pre}(x,y) & = & S(x,y) \circ Psf_S(x,y,u,v) - T(x,y) \otimes K(u,v,x,y) \nonumber \\ 
D'_{Pre}(x,y) & = & D_{Pre}(x,y). \nonumber 
\end{eqnarray}
By pre--smoothing the image that we are matching the template to --
$S(x,y) \circ Psf_S(x,y,u,v)$ -- we more frequently avoid the case
of psf--matching {\it deconvolution}, where the FWHM of $T$ is
narrower than $S$.  This pre--filtering increases the FWHM of the
image by $\sqrt{2}$.  The Psf matching kernels will need to be
correspondingly larger to account for the larger Psfs.  We are able to
run detection directly on this image.

In the second, which is the classical postfiltering implementation of
the technique, we create a difference image that is then
post--filtered with the Psf for detection:
\begin{eqnarray}
D_{Post}(x,y) & = & S(x,y) - T(x,y) \otimes K(u,v,x,y) \nonumber \\ 
D'_{Post}(x,y) & = & D_{Post}(x,y) \circ Psf_S(x,y,u,v).  \nonumber 
\end{eqnarray}

@@

Images $S(x,y)$ (optionally pre-filtered with $Psf_S(x,y,u,v)$) and
$T(x,y)$ are sent to ImagePsfMatchTask, whose purpose is to fit for
$K(u,v,x,y)$ and to produce $D(x,y)$.  The first step is to use the
Psf Gaussian sigmas to define the {\it sizes} of the Gaussians shapes
in $K_i(u,v)$.  As mentioned above, the main width is chosen to be
$\sqrt{\sigma_S^2 - \sigma_T^2}$.  We investigate variations on the
configuration where we use one Gaussian with the above width, and
where we use three Gaussians, the central of which has this width and
the others are a factor of $\beta = 2.0$ smaller and larger.  We
modify each of the Gaussians by a set of Laguerre polynomials of a
given order.  The smallest Gaussian is modified to the highest order,
with the others modified by order//2.  Figure~\ref{basis} provides an
example basis set with this configuration.  The total number of shapes
in the basis are: $\sum_i^{\rm nGauss} ({\rm degGauss}_i+1)\times({\rm
  degGauss}_i+2)/2$ \footnote{ A notable exception to this algorithm
  is found in the case where $\sigma_T > \sigma_S$, i.e. a sharpening
  must be done to the template to match it to the science image.  This
  is the case of deconvolution.  A theoretical prescription for basis
  design in this case is given in \cite{0266-5611-26-8-085002}.  In
  this algorithm, nGauss is {\it always} fixed to value 3, and
  degGauss is always fixed to 3.  The widths of the Gaussians are
  determined through the algorithm specified in
  \cite{0266-5611-26-8-085002}, using as inputs the sequence of
  Gaussians that we would have used to match a Gaussian of width
  $\sigma_S$ to $\sigma_T$ (i.e. as if we would have convolved the
  science image and not the template image).  There are a variety of
  reasons that actually operating on the science image is undesirable,
  including: the science image is lower S/N than the template; the
  science image is likely to have masked defects, including cosmic
  rays and bad pixels, whose reach would be expanded due to the
  convolution; unmasked features such as faint cosmic rays are more
  likely to look Psf--like after convolution (cosmic rays will likely
  have been sigma clipped out of the coadd).  However, it is not clear
  that deconvolving the template image is in principle any better.  A
  far better option, which we explore here, is to pre--filter the
  science image, which will in the vast majority cases provide a
  larger Psf (larger by $\sqrt{2}$) to match the template to.  This is
  the case for v6866601 in this study, where the postfiltered analysis
  yields deconvolution of the template, while the prefiltered analysis
  does not.  }.
%
The dimensions of the Psf matching kernel are
chosen to be 6 times the largest Gaussian width, with a minimum
kernelSize of 21x21 pixels.  For kernels that are significantly
smaller than this, the Gaussians have significant (non--zero) power at
the kernel boundaries, leading to square systematic artifacts at the
scale of the kernel in the difference images.

@@

For each source $j$ returned by DiaCatalogSourceSelector, we create a
{\tt KernelCandidate} instance, which holds substamps of $T_j(x,y)$
and $S_j(x,y)$ centered on the source.  These will be used to fit for
a local solution $K(u,v)$; the ensemble of local solutions at each
source's $x,y$ will be used to fit for the full spatial solution
$K(u,v,x,y)$.  We emphasize that the dimensions of the substamps are
important.  Note from equation~\ref{eq-soln} that the coefficients
$a_i$ are derived from a convolution of the template substamp with the
kernel basis functions $K_i$.  This convolution means that
kernelSize//2 pixels along each substamp edge are rendered unusable.
To maintain a significant number of pixels in $C_i$, we set the {\tt
  KernelCandidate} dimensions to be twice that of the kernel (i.e. the
number of pixels remaining in $C_i$ is equal to the number of pixels
in the kernel, which is fewer than the number of bases $i$).  For each
{\tt KernelCandidate}, we solve for $K(u,v)$ and create a local
difference image $D_j(x,y)$.  We evaluate several statistics on this
local difference image, normalized by the square root of its variance,
which puts the pixels in units of sigma.  We measure the mean sigma,
the RMS of this distribution, the $\chi^2$ of these pixels normalized
by the degrees of freedom (number of pixels in $D_j(x,y)$ minus the
number of kernel bases), and the mean squared error.  These are
referred to as the LOCAL kernel metrics, and reflect how appropriate
the chosen kernel basis is.

@@

Finally, we run DipoleMeasurementTask on the detections to create {\tt
  DiaSources}.  DipoleMeasurementTask is an extension of
SourceMeasurementTask that includes specific dipole measurement
algorithms.  Measurement is necessarily different in the pre--filtered
and post--filtered data.  In particular, in post--filtering, detection
happens on the filtered maximum likelihood image and source
measurement on the associated pixels in the unfiltered difference
image.  In pre--filtering, the latter does not exist, and thus
detection and measurement happen on the same image.  This restricts
the types of measurements that may be made on the sources.  For
example, the Psf flux is merely the fitted amplitude of the peak in
the image, and the Psf centroid its center.  While these measurements
are trivial to implement, additional work needs to be done on how to
implement the remainder of the measurement suite in a manner
comparable to the measurements on post--filtered data.

All {\tt DiaSources} are associated with the {\tt calexp's} sources
and with the reference catalog, with a permissive matching radius of
3''.  {\tt DiaSources} that match with {\tt sources} may be true
residuals around stars {\it or} true false detections that are also
found in the {\tt calexp} SourceDetectionTask.  For this reason, we
use the associations with the reference catalog objects as the
assessment if a given {\tt DiaSource} matches with an input object or
whether it is an orphan (e.g. a statistical fluctuation in the
background).  In addition, all {\tt KernelCandidates} are persisted
with the above {\tt LOCAL} and {\tt SPATIAL} metrics.  Repositories
are accessed with a specially written RepositoryIterator that
aggregates {\tt KernelCandidates} from a given source across multiple
output repositories.

@@

We ran several pre--productions to tune the following: the dimensions
of the kernels and {\tt KernelCandidates}; the widths of the Gaussians
(sigGauss) based on the FWHM of the input images; and the scaling of
the Gaussians ($\beta$) in the basis sequence.
Figure~\ref{fig_galleryBad} provides a rogue's gallery of image
subtraction failures, centered on {\tt KernelCandidates}, and outlines
reasons for each failure.  Figure~\ref{fig_galleryGood} provides a
gallery showing what the results from a successful image subtraction
appear like, for postfiltering on the {\it top} and prefiltering on
teh {\it bottom}.

The main production runs included: an outer loop over the use of
pre--filtering or post--filtering for detection; variations of the
complexity of the Kernel basis, including the number of Gaussians
(nGauss = 1,3) and the degree of the modifying Laguerre polynomials
(degGauss = 1..6); and the spatial order of the Chebyshev polynomials
on $a_i(x,y)$ (spatialKernelOrder = 1..6).  Finally, we ran several
post--productions focused on perturbing the solutions that yielded the
fewest numbers of DiaSources along specific dimensions.  This
included: the RMS of the WCS solution used by RegisterTask; including
a DC offset in the source positions prior to RegisterTask, to
determine the ability of the kernel basis to accommodate astrometric
misregistration; the size of the SpatialCells used in the spatial
interpolation, which effectively regulates the numbers of constraints
that go into the spatial fit; and modifying the detection thresholds
to determine how the number of false detections scales with threshold.

@@

In order to investigate the bias--variance trade--off in the overall
fit, we calculate the MSE of each {\tt KernelCandidate's} difference
image derived from the evaluation of the spatial model at that
location.  We define the bias as $\left| data - model \right|$, the
variance as $\left| (data - model)^2 \right|$, and the MSE as {\tt
  bias$^2$ + variance}.  In this context, the bias is the mean of the
difference image, and the variance is the mean square of the
difference image.  In both cases, we normalize by the square root of
the variance so that pixels are in units of sigma.

@@

After running an ensemble of 144 pipeline configurations, we
aggregated the numbers of false detections.  We report the results for
the each of the 3 visits, and for pre-- and postfiltering for
detection.  We cut on {\tt flags.pixel.edge} to remove spurious
detections on the immediate border of the image; other than that we do
no other cuts other than requiring that the {\tt centroid.sdss} is
finite.  We separated the DiaSources that match with the reference
catalog (3'' matching radius) from those that don't (referred to as
orphans) to investigate the origins of the detections.  We found no
significant correlation of false detections with sensor within the
raft; therefore we report the aggregate results for all 9 sensors in
the analyses that follow.

@@

The numbers of false detections in each best case are given in
Table~\ref{tab-bestfp10}.  We undertook a manual inspection of all
difference images to associate the DiaSources with Sources in the
images, and also use a 3'' association with the input source catalog
to define matches and orphans.  Associations are listed in the
N$_{Match}$ column of Table~\ref{tab-bestfp10}, with the unmatched
false positives in the N$_{Orphan}$ column.  The former are likely to
come from systematics in the subtractions of stars, while the latter
would be due to fluctuations in the background.  We note 2 effects
that will impact N$_{Match}$.  The first is that, with a 3'' matching
radius, 4\% of random fluctuations will by--chance associate with a
Source.  Assuming that N$_{Orphan}$ represents 96\% of the background
fluctuation population, we estimate the number of fluctuations from
that same population that would have associated with Sources as
N$_{Random}$.  Finally, we count all matches that occur within 100
pixels of the image boundary, where we might expect the spatial
extrapolation to yield systematic residuals, as N$_{D<100}$.  To the
extent that (N$_{Random}$ + N$_{D<100}$) $<$ N$_{Match}$, we have a
small residual set of systematic false detections in our imaging
ensemble.  Aside from the deconvolution run, this is at the level of
10 total false detections over 9 sensors in 5 production runs, or 1
every 4--5 images.

@@

After finding the configurations that yielded the minimal numbers of
false detections, we then perturbed the solutions to examine how the
numbers of false detections scaled with different effects.  This
included, in order of importance for this production: the difference
image detection threshold; the number of {\tt KernelCandidates} going
into the spatial model; and astrometric registration errors.  We also
undertook a theoretical analysis of the number of false detections
expected in a 4k x 4k image as a function of detection threshold, and
examine the requirements this sets on our ability to model the
variance and backgrounds in the images.  We compare the empirical
variation of false detections with detection threshold in our data,
and show that our results are consistent with this, indicating that we
reached the systematic floor in these data.

@@

Because these images will have gone through multiple convolutions
before reaching the detection stage, it is expected that the
propagated per-pixel variance will not provide an exact representation
of the empirical variance in the image planes \citep{Price-Stacking}.
Because the astrometric resampling uses a warping kernel that is
designed to preserve the noise properties of the images, we expect
that the kernel convolution and Psf--filtering will be the largest
sources of error.

@@

We further investigate the variance properties of the images that are
input to the detection stage, i.e. the difference image itself in the
case of prefiltering, and the difference image correlated with its Psf
in postfiltering.  These results are summarized in
Table~\ref{tab-variance2}.  We find that, empirically, the images have
very similar noise properties at the detection stage, with the
deconvolution visit having larger variance by $1\%$ but the other
visits having essentially equal noise properties, with a bias towards
the postfiltered data having slightly higher variance (while the RMS
is small, all values are greater than 1.0).  The medians of the
variance planes tell a different story though.  For the deconvolution
visit, the median per--pixel variance is $70\%$ higher.  While this
will suppress the detections of false positives, it will also
significantly suppress the detection efficiency for true variability.
For the other visits, the prefiltered per--pixel variance is
relatively {\it higher} by $3-4\%$.  This will decrease the
sensitivity of the prefiltered data to false positives; as shown
above, the rate is indeed smaller using the prefiltering pipeline.
This analysis demonstrates that this is primarily due to an {\it
  underestimate} of the per--pixel variance in the postfilter
processing: while the prefiltered variance is a 1\% underestiamte, the
postfiltered varaince is a 5\% underestimate.  As outlined by
\cite{Price-Stacking}, this may be due to differences in pixel
covariance when undergoing multiple convolutions, which are not
currently tracked.  We have not yet definitively demonstrated that
this is the case here.

@@

\item When deconvolving, the postfiltered variance is 70\% {\it
  higher} at the time of detection compared to the prefiltered data
  (Table~\ref{tab-variance2}).  This is consistent with the
  understanding that deconvolution increases high frequency noise in
  the images, and provides a {\it lower} detection efficiency in
  deconvolved data.  Prefiltering is clearly preferred in the case of
  significant deconvolution (v6866601; FWHM of 4.9 pixels matches to a
  FWHM of 3.5 pixels).

\item The postfilter pipeline can produce difference images with a
  minor deconvolution (v6866602; FWHM of 4.9 pixels matched to 4.9
  smoothed (v6866603; FWHM of 4.9 pixels matched to 6.7 pixels).

@@

\begin{figure}
\includegraphics[width=0.8\textwidth]{figures/shift.eps} \\
\caption{Figure illustrating the ability of the kernel basis to
  compensate, locally, for the effects of misregistration.  Each
  element of the grid shows a single {\tt KernelCandidate} local
  difference image (visit=6866603, doPreFilter=True).  Columns
  indicate the Gaussian sigma of the basis, which is designed to have
  one Gaussian with the prescribed width, modified to fourth order.
  Rows indicate the amplitude of the astrometric offset that was
  introduced into the Source positions before image-to-image
  registration.  In the {\it top} row, we see that even with a small
  astrometric offset, the sigma=2 basis is unable to produce a quality
  subtraction, because the width is inappropriate for matching the two
  input Psfs (the theoretical Gaussian sigma here is 3.4 pixels).  At
  sigma=4 the basis is able to produce a quality subtraction, and by
  sigma=5 the basis is too large to match the Psfs.  In {\it row 4},
  we show that even for an astrometric offset of 4 pixels (0.8''), the
  sigma=4 basis can produce a reasonable difference image without
  structured residuals.  However, by {\it row 5}, with an offset of 6
  pixels, none of the Gaussians provide a successful local difference
  image.  The ability of the basis to compensate for bulk astrometric
  offsets is a function of the kernel width, and the kernel width is
  itself a function of the two input Psf FWHMs.  Thus there is a
  complicated dependence of the ability of the kernel to shift the
  centroids of stars.  Roughly, the basis may compensate for local
  astrometric residuals at a scale up to $\sqrt{\sigma_S^2 -
    \sigma_T^2}$.  }
\label{kernel_offsets}
\end{figure}
